---
title: "Solution to Weight Lifting Problem"
author: "Akash Gupta"
date: "July 9, 2016"
output: html_document
---


### first of all we load necessary files namely training and testing , which were ###provided to us.


```{r}
training = read.csv("pml-training.csv", stringsAsFactors = F)
testing = read.csv("pml-testing.csv", stringsAsFactors = F)
```

### Now we have to clean the data as many of the variables have a lot of missing ###values.
### First we will remove any variable which has more than 60% of data missing and ###then we will remove all non numeric variables.
### However we will retain variable "classe" since we have to predict it.

```{r}
library(caret)
library(randomForest)
counter = 0
newData = data.frame(matrix(nrow=length(training[,1]), ncol=7))
for(i in 1:length(training)){
  y = sum(is.na(training[,i]))
  n = y/(length(training[,1]))
  if(n < .6){
    counter = counter + 1
    newData[,counter] = training[,i]
    colnames(newData)[counter] = colnames(training)[i]
  }
}
numbers = sapply(newData, is.numeric)
new1 = cbind(newData[,numbers], classe = newData$classe)
new1 = new1[,5:length(new1)] 
```

### Since we will test our final model on testing dataset and we will not want to ###touch it before we come up with our final model,
### we will split our current dataset which is new1 in two parts, Train and Test ###with 60% of data in train and rest in Test. 

```{r}
set.seed(1000)
inTrain = createDataPartition(y=new1$classe, p=0.6, list=FALSE)
Train = new1[inTrain,] 
Test = new1[-inTrain,]
```

### Now since we have got our final training set (Train) we will fit our model on ###it & will use random forest as classification method.
### We are choosing random forest as it gave highest accuracy among other models.
### We will use train control method cv with 4 partition and will allow parallel ###processing for fast results.

```{r}
set.seed(1000)
modelRF = randomForest(classe~., data = Train, trControl = trainControl(method = "cv", number = 4, allowParallel = TRUE))
```

### Now we classify this model on Test set.
```{r}
prediction = predict(modelRF, Test)
confusionMatrix(Test$classe, prediction)
```

### But since we have to finally predict on the testing file , we do it here.
### But to predict here on testing we have to process it as training file

```{r}
Needed = colnames(Train)
Needed[53] = "problem_id"
testFinal = testing[, Needed]
predictionTest = predict(modelRF, testFinal)
```

### We will also want to print the prediction vector for testing file.

```{r}
print(as.vector(predictionTest))
```






